Generating Interactive WebSocket Applications in TypeScript

### 2020-03-09 PLACES_2020/casestudy.tex
More proofreading

Note that, by TypeScript's structural typing 
\cite{TypeScriptSpec}, replacing
\texttt{handleP2Move} on \cref{lineupdate} with a recursive 
occurrence of \texttt{handleP1Move} would be
type-correct -- this allows for better code reuse as opposed to 
defining
additional abstractions to work around the limitations of nominal 
typing in
\cite{Hybrid2016}.
There is also full type erasure when transpiling to JavaScript to 
run the server code, so state space explosion is not a 
runtime consideration.

@@

For the sake of code reuse, \cite{NoughtsAndCrosses}
uses \textit{higher-order components} (HOC) to build the correct state
implementations depending on which player the user chooses to be.
We leverage the \textit{Redux} \cite{Redux} state management library to keep track of game
state, thus showing the flexibility of our library design in being
interoperable with other libraries and idiomatic JavaScript practices.
% REVIEW: tight-coupling?
Our approach encourages the separation of concerns between the 
communication logic and program logic -- the generated session runtime
keeps track of the state of the EFSM to ensure protocol conformance by 
construction, whilst \textit{Redux} solely manages our game state.

>>>>

Note that, by TypeScript's structural typing
\cite{TypeScriptSpec}, replacing
\texttt{handleP2Move} on \cref{lineupdate} with a recursive
occurrence of \texttt{handleP1Move} would be
type-correct --- this allows for better code reuse as opposed to
defining
additional abstractions to work around the limitations of nominal
typing in
\cite{Hybrid2016}.
There is also full type erasure when transpiling to JavaScript to
run the
server code, so the types defined in TypeScript will not appear in the
JavaScript after typechecking.
This means state space explosion is not a runtime consideration.

@@

For the sake of code reuse, \cite{NoughtsAndCrosses}
uses \textit{higher-order components} (HOC) to build the correct state
implementations depending on which player the user chooses to be.
We leverage the \textit{Redux} \cite{Redux} state management library to keep
track of the game state, thus showing the flexibility of our library design in
being interoperable with other libraries and idiomatic JavaScript practices.
% REVIEW: tight-coupling?
Our approach encourages the separation of concerns between the
communication logic and program logic --- the generated session runtime
keeps track of the state of the EFSM to ensure protocol conformance by
construction, whilst \textit{Redux} solely manages our game state.

### 2020-03-09 PLACES_2020/conclusion.tex
More proofreading

\section{Conclusion and Future Work}
We have presented a MPST-based framework for developing full-stack interactive
TypeScript applications with WebSocket communications that conform to a
protocol, statically providing linear channel usage guarantees and affine
channel usage guarantees for back-end and front-end targets respectively.

>>>>

\section{Conclusion and Future Work}
We have presented a MPST-based framework for developing full-stack interactive
TypeScript applications with WebSocket communications.
The implementation conforms to a specified
protocol, statically providing linear channel usage guarantees and affine
channel usage guarantees for back-end and front-end targets respectively.

### 2020-03-09 PLACES_2020/related.tex
More proofreading

\paragraph{Endpoint API Generation}
Neykova and Yoshida targeted Python applications and the generation of runtime
monitors \cite{Python2017} to dynamically verify communication patterns.
Whilst the same approach could be applied to JavaScript, we can provide more
static guarantees with TypeScript's gradual typing system and compiler. Scribble-Java \cite{Hybrid2016} proposed to encode the EFSM
states and transitions as classes and instance methods respectively, with
behavioural typing achieved statically by the type system and channel linearity
guarantees achieved dynamically since channels are exposed and
aliasing is not monitored.
% REVIEW: could the approach of the present paper by applied to Java as well? discuss?
Scribble-Java can generate callback-style APIs similar to the approach we 
present, but this approach is arguably less idiomatic for Java developers.

\paragraph{Session Types in Web Development}
King et al. \cite{PureScript2019} targeted web development in PureScript using the
\textit{Concur UI} framework and proposed a type-level encoding of EFSMs as
multi-parameter type classes.
However, it presents a trade-off between achieving static linearity guarantees
from the type-level EFSM encoding under the expressive type system and
providing an intuitive development experience to developers, especially given
the prevalence of JavaScript and TypeScript applications in industry. Fowler \cite{MVU2019} focused on applying binary session types in front-end web
development and presented approaches that tackle the challenge of guaranteeing
linearity in the event-driven environment.

>>>>

\paragraph{Endpoint API Generation}
Neykova and Yoshida targeted Python applications and the generation of runtime
monitors \cite{Python2017} to dynamically verify communication patterns.
Whilst the same approach could be applied to JavaScript, we can provide more
static guarantees with TypeScript's gradual typing system.
Scribble-Java \cite{Hybrid2016} proposed to encode the EFSM
states and transitions as classes and instance methods respectively, with
behavioural typing achieved statically by the type system and channel linearity
guarantees achieved dynamically since channels are exposed and
aliasing is not monitored.
% REVIEW: could the approach of the present paper by applied to Java as well? discuss?
Scribble-Java can generate callback-style APIs similar to the approach we
present, but this approach is arguably less idiomatic for Java developers.

\paragraph{Session Types in Web Development}
King et al.\ \cite{PureScript2019} targeted web development in PureScript using the
\textit{Concur UI} framework and proposed a type-level encoding of EFSMs as
multi-parameter type classes.
However, it presents a trade-off between achieving static linearity guarantees
from the type-level EFSM encoding under the expressive type system and
providing an intuitive development experience to developers, especially given
the prevalence of JavaScript and TypeScript applications in industry.
Fowler \cite{MVU2019} focused on applying binary session types in front-end web
development and presented approaches that tackle the challenge of guaranteeing
linearity in the event-driven environment, whereas our work is applicable to
multiparty scenarios.
